---
title       : Regression to the Mean
subtitle    : Children's Heights from Galton
author      : Chris Knutson
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [bootstrap]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Regression to the Mean

### Key Principal

- Things tend to even out over time
- If you have an extreme measurement, your next measurement is likely to be less extreme

---

## Galton Data

Summary: Galton data shows that children's heights tend to regress to the mean

- Tall parents have tall children
- Short parents have short children
- Very tall parents have children shorter than themselves
- Very short parents have children taller than themselves

---

## Illustration via Linear Regression

```{r, echo=FALSE, results='hide'}
library(UsingR)
data(galton)

childmu <- mean(galton$child)

parent <- 72
lmfit <- lm(child ~ parent, data = galton)
pred <- predict(lmfit, newdata = data.frame(parent))
```
```{r, echo=FALSE}
hist(
     galton$child, 
     xlab = "Children's Heights", 
     xlim = c(60,75),
     col = 'lightblue', 
     main='Predicted Child Height Relative to Mean Child Height'
)
lines(c(parent, parent), c(0,200), col = 'blue', lwd=5)
lines(c(childmu, childmu), c(0,200), col = 'green', lwd=5)
lines(c(pred, pred), c(0, 200), col='red', lwd=5)
legend("topright", c("Parent height", "Child Mean", "Predicted Child Height"), 
       col=c("blue", "green", "red"), lwd=10)
```

---

## Try it Out!

Test predictions based on a linear model of the Galton data using my ShinyApp:

- [Galton Child Height Regression](https://canuteson.shinyapps.io/galton_regression/)


